### 2010-03-11 papers/icfp-2010-desc/paper_desc.tex
Desc@ICFP: fix \cite{} which was preventing compilation on ancient version of pdflatex.
Desc@ICFP: transform my TeX comments into readable (haha) notes.
Desc@ICFP: more philosophical questions

%%! Check Martin-Lof reference

For our sigma-of-sigmas to be known, we crucially need the ability to
name them. To this end, we will use a standard technique in
dependently-typed programming: we construct an ad-hoc universe. This
technique dates back to Martin-L\"of definition of type
theory~\cite{martin-lof:itt}. Since then, it has been fruitful as a
programming technique~\cite{}. We refer the reader to Agda's
tutorial~\cite{norell:agda-tutorial} for a pedagogical presentation of
universe construction.

@@

@@

@@

@@

%%! Strictly positive types anyone?

@@

@@

@@

%% I'm kind of shafted (haha) by this:
%%     - We have not yet introduced tagged descriptions
%%     - Tagged description in a Mu, that just doesn't work...
\[
\Rule{\forall e_i \in \vec{e}, \propag{\push{e_i}
                                            {\descop{f\: i}{(\Mu{(\DSigma{(\Enum\: b)}{f})})}}}
                                      {e'_i}}
     {\propag{\push{C\: \vec{e}}{\Mu{(\DSigma{(\Enum\: b)}{f})}}}
             {\Con{(\etag{C}, \widetilde{e'})}}}\;\etag{C} \in b
\]

>>>>

\note{Check Martin-Lof reference}

For our sigma-of-sigmas to be known, we crucially need the ability to
name them. To this end, we will use a standard technique in
dependently-typed programming: we construct an ad-hoc universe. This
technique dates back to Martin-L\"of definition of type
theory~\cite{martin-lof:itt}. Since then, it has been fruitful as a
programming technique~\cite{who?} \note{I would need some other
  example of universe construction for practical purposes}. We refer
the reader to Agda's tutorial~\cite{norell:agda-tutorial} for a
pedagogical presentation of universe construction.

@@

\note{Is that really a motivating motivation for the usage of universe
  construction? }

@@

\note{We have talked about a tagged description based on $\List (\UId
  \times \Desc)$. It makes the free monad easier to construct but,
  imho, considerably clutters anything else. I've reverted back to a
  more sigma-friendly presentation.}

@@

\note{I think we need a function mapping a $\TagDesc$ to the
  corresponding $\Desc$. This function should go here.}

@@

\note{ Strictly positive types anyone? }

@@

\note{ I'm afraid that this section is killed by my categorical gibberish. }

@@

\note{Shall we give the code of $\All$?}

@@

\note{ This is rather drafty }
\[
\Rule{\forall e_i \in \vec{e}, \propag{\push{e_i}
                                            {\descop{f\: i}{(\Mu{(\DSigma{(\Enum\: b)}{f})})}}}
                                      {e'_i}}
     {\propag{\push{C\: \vec{e}}{\Mu{(\DSigma{(\Enum\: b)}{f})}}}
             {\Con{(\etag{C}, \widetilde{e'})}}}\;\etag{C} \in b
\]

### 2010-03-11 papers/icfp-2010-desc/paper_desc_levitation.tex
Desc@ICFP: finish IDesc section.
Desc@ICFP: transform my TeX comments into readable (haha) notes.
Desc@ICFP: more philosophical questions

%%! Maybe a bit of perspective would be good
%%    <- Everything stands up with the Mu
%%    <- Induction principle defined over Mu-things

@@

Consequently, the examples developped above can be straightforwardly
translated into this new presentation. In the following, we adopt this
last version as our de facto universe of inductive types. In
particular, we are going to demonstrate several instances of generic
programming.

%%! Shall we talk about the Type Theory being Desc Zero? or such story?

@@

%%! Update to the factorized presentation

@@

\subsection{The generic Free Monad}

@@

%%! Recall monadic structure of this object

@@

%%! This is not quite Mu of a tagged description

@@

%%! Triple check. That doesn't work, right?

>>>>

\note{
Maybe a bit of perspective would be good: 
Everything stands up with the Mu.
Induction principle defined over Mu-things.
}

@@

Consequently, the examples developped above can be straightforwardly
translated into this new presentation. \note{We might want to re-do
  the Nat example here, as we encode once again in $\IDesc$} In the
following, we adopt this last version as our de facto universe of
inductive types. In particular, we are going to demonstrate several
instances of generic programming.

\note{Shall we talk about the Type Theory being Desc Zero? or such story?}

@@

\note{Update to the factorized presentation}

@@

\subsection{The generic Free Monad}
\label{sec:desc-free-monad}

@@

\note{Recall monadic structure of this object}

@@

\note{This is not quite Mu of a tagged description}

@@

\note{ Triple check. That doesn't work, right? }

### 2010-03-11 papers/icfp-2010-desc/paper_type_theory.tex
Desc@ICFP: transform my TeX comments into readable (haha) notes.
Desc@ICFP: more philosophical questions

@@

@@

@@

%%! This might deserve a more thorough explanation, with snippet of code.

@@

@@

In this section, we have developped a type propagation system, based
on bidirectional type-checking. Using bidirectional type-checking as a
boilerplate scrapper is a well-known technique~\cite{who?}. In our case,
we have shown how to instrument bidirectionality to rationalize the
expressivity of dependent types. We have illustrated our approach with
finite sets. We have successfully abstracted away the low-level
presentation of finite sets, offering a convenient syntax instead.

>>>>

\note{Need some care here. Expansion would be good too. I wanted to
  carry the intuition that we are not the bad guys with a
  non-terminating type-checker.}

@@

\note{Shall we talk about W-types encoding?}

@@

\note{Did I got the coding presentation vs. coding representation
  story right?}

@@

\note{ This might deserve a more thorough explanation, with snippet of
       code. }

@@

\note{And suddenly, type-directed pretty printing comes into the
  picture. Surely, it should ``just'' be the inverse of
  elaboration. But is it? And how do we present it?}

@@

In this section, we have developped a type propagation system, based
on bidirectional type-checking. Using bidirectional type-checking as a
boilerplate scrapper is a well-known technique~\cite{who?} \note{I
  would need some citation of people using bidirectional presentation
  of their type theory}. In our case, we have shown how to instrument
bidirectionality to rationalize the expressivity of dependent
types. We have illustrated our approach with finite sets. We have
successfully abstracted away the low-level presentation of finite
sets, offering a convenient syntax instead.
